Staying Hydrated: The Molecular Journey of Gaseous Sulfur Dioxide to a Water Surface

### 2012-03-27 resubmission/ResponseToReviews.tex

Journal: The Journal of Physical Chemistry 
Manuscript ID: jp-2011-11804m 
Title: "Staying Hydrated: The Molecular Journey of Gaseous Sulfur Dioxide to a 
Water Surface" 
Author(s): Shamay, Eric; Richmond, Geraldine

Dear Prof. Richmond:

Thank you for submitting your manuscript for publication in The Journal of Physical Chemistry. It has been examined by expert reviewers who have concluded that the work is of potential interest to the readership of The Journal of Physical Chemistry; however, it appears that a major revision, possibly followed by further reviewer evaluation, will be needed prior to its further consideration for publication. Please see the enclosed reviewers' reports for details regarding the requested changes and/or additions.

Additionally, we have noticed that your manuscript referenced above is missing the Table of Contents graphic. Please see the submission requirements for the Journal of Physical Chemistry TOC Graphic at http://pubs.acs.org/paragonplus/submission/jpchax/jpchax_authguide.pdf. 
Note that the graphic for the TOC entry should be in the form of a structure, graphical image, photograph, reaction scheme, or equation. The Author must submit the graphic in the actual size to be used for the TOC, fitting in an area 1.375 in. high and 3.5 in. wide (3.6 cm Ã— 8.9 cm). The allowed file formats for TOC graphics are TIFF and PDF. Please submit a revision of this manuscript including the TOC graphic as the final page of your Article manuscript revision submission. Separate TOC graphics are not allowed.

We ask that you submit your revision no later than 29-Apr-2012.

When submitting your revised manuscript through ACS Paragon Plus, you will be able to respond to the reviewers' comments in the text box provided or by attaching a file containing your detailed responses outlining the changes made and explaining your reasons for disagreeing with any suggestions you choose not to follow.

To Revise Your Manuscript on the Web:

To revise your manuscript, log into ACS Paragon Plus at http://paragonplus.acs.org/login and select "My Authoring Activity". There you will find your manuscript title listed under "Revisions Requested by Editorial Office." Your original files are available to you when you upload your revised manuscript. If you are replacing files, please remove the old version of the file from the manuscript before uploading the new file.

Thank you for considering The Journal of Physical Chemistry as a forum for the publication of your work.

With sincere regards,

Prof. Pavel Jungwirth 
Senior Editor 
The Journal of Physical Chemistry 
Journal Fax: 202-354-5425 
Journal Email: jungwirth-office@jpc.acs.org

Reviewer: 1 
-----------

1,2,3. The authors have taken into consideration the entirety of all the reviewer suggestions regarding functional and basis sets, cutoffs, timesteps, and temperature that affect the simulations. The changes are reflected in the text, and meet the standards suggested in the reviews. All the simulation data has been re-analyzed, and the results are all presented from the new simulations.
% 1. The computational parameters chosen as well as the authorsʼ claim that “these computational parameters were verified to yield a reasonable description of bulk room temperature water …” runs counter to published results.

%2. Given that the system contains water and thus the highest vibrational modes are ~3700 cm-1, it is hard to fathom a timestep of “1 fs” can be used to accurately integrate the equation of motions without energy drift. If this is truly the case, it might behoove the authors to provide some supplemental materials showing the conserved quantity as a function of time for both the NVT and NVE simulations provided the system is not glassy and can sample all the possible conformations.

%3. The author stated that DZVP basis set was chosen, it was never explained why the authors believe that such a basis set is adequate. In fact, work by VandeVondele et al shows that at least a TZV2P basis set with a cutoff of 280 Ry was necessary in order to obtain forces comparable to other more established codes such as CPMD. In addition, works by Siepmann et al on vapor liquid coexistence curves hints that the standard TZV2P basis set with a cutoff of 280 Ry might still be inadequate. Last, what was the convergence criteria used for the SCF?

%4a. The cold system temperature was set to 210K in the resimulated systems. This much larger separation of temperatures is more in line with the reviewer's concerns regarding thermostatting. 
4. Following recent literatures on simulations of liquid water using potentials derived from density functional theory (albeit some with different codes as well as different methods), one firm conclusions can be made is that water near ambient conditions are more glassy than liquid. One way to overcome this is through massive thermostating which in addition also ensures equipartition of the kinetic energy in all degrees of freedom. This raises two issues.  a. First, as the production run was obtained in the microcanonical (NVE) ensemble and the two target temperature is only separated by 27 degrees, can the authors show that temperature for the subsystem of interest is really the targeted temperature thus fulfilling the equipartition theorem.  b. And can the authors show that indeed, the water part of the simulated system is behaving as a liquid and not conformationally trap in a glassy state?

5. “Nose” should be “Nosé”. In addition, the coupling frequency, when not chosen appropriately can cause energy drifts, was not specified.

6. The text has been updated to reflect the equilibration period of 1 ps.
%6. The authors indicate the system was equilibrated for “1 ns”. This just does not seem correct unless the equilibration was done using a cheaper potential other than DFT.

7. From the RDF, it can be seen that both hot and cold simulations converges to 1.0 as distance increases. But if the sulfur dioxide gas molecule is at the liquid-vapor interface, shouldnʼt this value approach something more like 0.5 instead of 1.0?

8. The inclusion of dispersion corrections were not explored for this simulation study. The goal of the project was to monitor specific bonding behaviors of SO2 using previously established methods, not as a study of the results from various constraints or changes to simulation parameters.
%8. Has the authors explore the ramifications of including dispersion correction in their calculation which might play a large in such a system?

Reviewer: 2 
-----------

Recommendation: This paper is probably publishable, but should be reviewed again in revised form before it is accepted.

Reviewer: 3
-----------

1. Because the re-simulated systems were done using B3LYP functional a benchmark comparison is no longer necessary. Small cluster studies of so2 with several waters have been performed using high-accuracy basis and functionals previously, and we do not set out to reproduce that work here, but rather to build on it. The authors believe that the current work with the new simulation parameters achieves a level of accuracy suggested by the reviewers, but that lengthy discussion of this fact is not requisite in this particular simulation study.
%1.	The predominant stabilization interaction between and SO2 molecule and a water molecule is due to dipoles. This essentially places the sulpher close to the water oxygen. This simple dipole interaction reasoning is consistent with everything the authors find. However, there is a component of this interaction that arises due to van der Waal’s, due to the sulpher lone pairs. In this regards BLYP has been used here, which is known not to capture such subtle interactions. Considering this, I would have liked to see at least a set of benchmarks for the functional and basis used in this work, with post Hartree-Fock MP2 and B3LYP used for comparison. The benchmark system could have involved a simple cluster of SO2 with a few water molecules. The absence of this comparison makes the results weak.

2.	The larger concentration of unbound SO2 at a lower temperature is counter-intuitive. (Fig. 4) One would have expected the opposite trend and no physical reasoning is provided to explain this result.

3. Several previous simulation studies have discussed the structure of water under similar simulated conditions. The focus of the work is on the structure of SO2 and its behavior near a water cluster surface. Thus, the authors have made the conscious decision to avoid lengthy discussion of the surrounding water's structure.
%3.	The structure of the surrounding water molecules is never discussed.

4.	It is stated that Quickstep is used. Assume this implies that the Car-Parrinello method is used, but it is not stated in the paper.

5.	The system includes 36 water molecules in a box. The effect of this restriction is important to discuss.

6. K-point integration was not used in the integration, as this was not addressed in similar previous studies. Also, the CP2K simulation package does not allow for this type of calculation. The authors do not believe this is detrmiental to the accuracy of the calculations.
%6.	Was k-point integration used in the calculation? If so how many. If only gamma-point was used, again there needs to be a discussion on the effectiveness of the calculation.

7. In our new simulations, the timestep has been reduced to 0.5 fs. The text has been updated to reflect this.
%7.	The time-step of 1fs appears to be too large for a Car-Parrinello simulation. Since NVE simulations are used, I would be interested in knowing how well the total system energy is conserved (which determines the extent to which the dynamics simulations are accurate).

% 8.	With respect to previous work on molecular computations using graphs, I would like to bring to the author’s attention the following work by O. Sinanoglu, which to my knowledge is one of the first such attempts: Oktay Sinanoglu J. Am. Chem. Soc., 1975, 97 (9), pp 2309–2320. 
8. The authors thank the reviewer for pointing out this previous study using graph theoretical techniques (Oktay Sinanoglu J. Am. Chem. Soc., 1975, 97 (9), pp 2309–2320). That study pertains to synthetic or catalytic pathways of organic systems, but can be generalized to very convoluted and complicated reaction pathways leading to very intricate graphical networks. The systems investigated in that work are much more complicated, and the graphs are designed to convey a great deal more information than in the present text. In our work we wish to investigate only the transitions between different coordination or cyclic-hydrate states, and are not interested in multi-reactant processes. A citation to the work has been added to the text.

### 2012-03-27 resubmission/Article.tex
Preparing the resubmission files. Working on responses to the reviewers.

Quantum MD techniques are the logical follow-up as they accurately reproduce the hydration geometry around the bound \suldiox~molecules, and allow us to examine in detail the specific bonding interactions that occur within the surface hydrates, and in the extended bonding further into the water.\cite{Baer2010} Two parallel studies are performed in this work; one at 300K, and one at the more atmospherically relevant cold 273K. This set of temperatures complements our most recent experimental studies that showed the binding of gaseous \suldiox~to a water surface is greatly enhanced at cold temperatures.\cite{Ota2011} Other experiments by our group developed the picture of \suldiox~adsorption, and showed that \suldiox~surface hydrate complexes form when a water surface is exposed to \suldiox~gas.\cite{Tarbuck2005,Tarbuck2006} Although conclusions regarding the specific nature of those complexes could only be inferred from the experiments, our current computational studies now provide us with insights about the specific microscopic geometries and behaviors of the hydrating complexes.

@@

As noted above, the optimized geometry of the gas phase \suldiox~hydrates suggests cyclic bonding structures. Geometry optimization shows the formation of these cyclic hydrate structures with two or three waters. A different story entirely has the potential to emerge when \suldiox~is placed in a dynamic environment such as in the course of MD simulations of an aqueous surface. Do the cyclic structures also form in the course of a dynamic bonding process on a simulated water surface, where extended hydrating structures influence \suldiox~and water behavior? To study the formation and behavior of cyclic hydrate structures we employ graph theoretical techniques on MD trajectory data. Previous use of graphs in molecular computations were applied to finding stable arrangements of water clusters, ice, hydrogen bonding, extracting topological molecular properties, and cyclic structure studies.\cite{Anick2002, Huber2007, Radhakrishnan1991, Shi2005, Garcia2004, McDonald1998}

@@

On-the-fly ab initio molecular dynamics simulations were performed with the QUICKSTEP package, which is an implementation of the Gaussian plane wave method using the Kohn-Sham formulation of density functional theory (DFT).\cite{VandeVondele2005} The Kohn-Sham orbitals are expanded using a linear combination of atom-centered Gaussian-type orbital functions. The electronic charge density was described using an auxiliary basis set of plane waves. Energies and forces from on-the-fly simulation sampling of the Born-Oppenheimer surface were calculated for each MD step using the Gaussian DZVP basis set, the exchange-correlation functional of Becke, Lee, Yang, and Parr (BLYP),\cite{LEE1988} and the atomic pseudo-potentials of the Goedecker, Teter, and Hutter type.\cite{Goedecker1996} A simulation timestep of 1 fs was used, with a Nose-Hoover thermostat set at 273K and 300K for the ``cold'' and ``hot'' simulations, respectively. These computational parameters were verified to yield a reasonable description of bulk room temperature water when simulating a neat-water system.

Initially, 10 equilibrated boxes of side-lengths 10.0\angs, with 36 randomly packed water molecules were used. Five of the boxes were used for each of the cold and hot simulations. A sulfur dioxide molecule was randomly placed onto the surface within 2.5\angs~of a water molecule centrally located above the waters in the z-axis. A copy of the initial system cubes were then expanded along one axis (z-axis) to 25\angs. The system energy was minimized through a geometry optimization. Subsequently, the system was equilibrated for 1 ns in canonical ensemble (NVT) conditions. Periodic boundaries were set on the two short axes to form an infinite slab. The equilibrated systems were then simulated for a further 20 ps in the microcanonical ensemble (NVE), with trajectory snapshots recorded every 1 fs. The initial 1 ns equilibration trajectory was not included in the final analysis. This simulation process resulted in 20,000 time steps of system trajectory for analysis in each of the hot and cold replicas of the system, for a total of 100,000 timesteps at each temperature.

@@

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{so2-coordinations-percents-small.png}
		\caption{The \suldiox~bonding coordinations occur with different frequencies at the surface under hot and cold temperature conditions. Shown above is the distribution of bonding coordinations of the cold (blue) and hot (red) \suldiox~over the entire set of trajectories. The values are the percentage of MD timesteps spent in the given bonding coordination.}
		\label{fig:bonding-coordinations}
	\end{center}
\end{figure}

@@

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{SO-cycle-breakdown-with-cartoons-small.png}
		\caption{Cyclic hydrate structures form throughout the MD simulations involving the \suldiox~and one or more water molecules. Shown here are the different cycles by number of water (up to 4), and their occurrence rate as a percentage of the total simulation time. The far right part of the plot shows the contributions of the two types of triple-water cycles (see Figure \ref{fig:type-3-varieties}). The cycles had to meet the ``minimally-SO'' bonding criteria (at least one bond to the sulfur and one oxygen bond on the \suldiox) to be used in the analysis.}
		\label{fig:cyclic-breakdown}
	\end{center}
\end{figure}

@@

We know that the \suldiox~bound to a water surface is most likely in the ``SO'' bonding coordination, and is also often taking part in some type of cyclic structure. With this in mind, it becomes interesting to ask: how long does a cyclic hydrate form before breaking to an acyclic hydrate structure? To answer the question of cyclic lifespan, a method was devised to define the lifetime of a cycle. For each MD trajectory, coordinate data was analyzed to determine if a cyclic hydrate structure was formed as described earlier in this manuscript. A timeline was then produced where each timestep was given a value of 1 or 0 depending on whether a \suldiox~hydrate bonding cycle was present or not, respectively. This resulted in a time-function, $C(t)$, similar in nature to a time-varying digital signal.

In an electro-mechanical system, a mechanical switch often outputs a noisy signal, full of transients or ``signal bounce'' before settling to a final value. This will appear as a rapid on-off cycling of the signal, and the problem is one of great concern in signal processing. Many mechanical, electrical, and software solutions have been devised to suppress the transients, or ``debounce'' the signal. Analogously, the hydrate bonding cycles formed in the simulated system often undergo a period of time during formation, or before dissolution, where the $C(t)$ function bounces before settling into a final value. The bounce in the function manifests itself in our statistics as a series of rapid switches between cycle presence and cycle absence. Physically, the bond lengths within the cycle (or the cycle-to-be) are fluctuating back and forth across our bond-length criteria. It is thus an artifact of our algorithmic determination of the presence or absence of a bond. Removing this artifact will allow us to examine the longer-time bonding behavior and this is accomplished as follows.

We wish to find a function based on $C(t)$ that eliminates the brief oscillations during transitions. The resulting function, $f(t)$, will only contain information of whether a hydrate bonding cycle is present or absent, and none of the noisy oscillations of the transitions between the two states. $f(t)$ may then be used to calculate statistics about the cycle lifespans. A representative portion of a cycle time-function, $C(t)$, is plotted for one of the simulated trajectories shown as the dashed black line in Figure \ref{fig:debouncing}. At the far left of the plot the function is in the ``no cycle'' state indicating that a bonding cycle has not been detected, and then switches ``on'' as a cycle is found later in time. The cyclic structure is very dynamic, constantly moving and distorting, so any of the bonds forming the cyclic bonding structure are liable to break and reform quickly. This bouncing between states is manifested in Figure \ref{fig:debouncing} as a series of sharp spikes in the $C(t)$ function lasting less than 10 fs each. $C(t)$ was smoothed using a moving gaussian window function with a 10 fs width, having the effect of disregarding cycle breaks or formations of less than 20 fs duration. The resulting smoothed function, $C_s(t)$, was then cutoff with the following criteria:

\[
  f(t) = \left\{ 
  \begin{array}{l l}
    0 & \quad C_s(t) < 0.2\\
    1 & \quad C_s(t) \geq 0.2\\
  \end{array} \right.
\]

where $f(t)$ is the debounced time-function that represents the lifespans of cycles. Figure \ref{fig:debouncing} shows the original time-function of cycle formation and breaking, $C(t)$ (dashed black), the smoothed function, $C_s(t)$ (red), and the final debounced function, $f(t)$ (green).

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{cycle-debouncing-small.png}
		\caption{The dynamic environment of the water surface causes rapid breaking and formation of bonds to the \suldiox, and to waters involved in the cyclic hydrating structures. A debouncing procedure was developed (described in the text) to account for the very brief ($<$ 20 fs) breaks to the cyclic structure. A time function, $C(t)$ (dashed black), shows if the bonding distance criteria are met for all members of the cyclic hydrate structure. A gaussian smoothing window was then used to produce $C_s(t)$ (red), which was then used with a cutoff criteria to give the debounced cycle formation time function, $f(t)$ (green).}
		\label{fig:debouncing}
	\end{center}
\end{figure}

The distribution of cycle lifetimes (i.e. contiguous spans of time spent with $f(t) = 1$) was determined. The distribution of cyclic lifespans is shown in Figure \ref{fig:cycle-lifespans} for the cold (blue) and hot (red) simulations. The most frequent lifespan for both temperature data sets lasts between 0-1 ps. This accounts for the vast majority of cyclic hydrates found (approximately 95\%) indicating that these structures are very transient, and are continuously forming and breaking for very short periods of time. Even with the debouncing procedure that would artificially increase the timespan spent either formed or broken, the nature of the water surface, and the very dynamic extended hydrogen bonding network, keeps many of the structures from lasting much longer than 1 ps. The difference between hot and cold systems in the $<1$ ps population is less than 2\%, with this trend extending to the longer lifespans as well. The inset of figure \ref{fig:cycle-lifespans} shows an expanded view of the region above 1 ps. All of the distribution shows only a $<1.5$\% difference between cold and hot cyclic lifetimes. Overall, the distribution shows that when cycles form, at both temperatures, they last a similar amount of time. Up to 3 ps, the cold temperature cycles show a very slight population increase above the hot temperature cycles. Above 4 ps, most of the cycles that form are found in the hot system. The 8 ps cycles are notable in that they form for just under half the length of one of the simulated trajectories.

We know from the transition frequency plots of Figure \ref{fig:coordination-transitions} that the bonding coordinations are switching frequently. The \suldiox~is likely forming and breaking bonds with waters external to the cyclic hydrate structures (i.e. not directly involved in the bonds of the cycle). For the longer-lived cycles, the external bonding to the \suldiox~may have little effect on the cyclic hydrate waters. However, any time the \suldiox~switches into an unbound, sulfur-only, or oxygen-only coordination (i.e. ``S'', ``SS'', ``O'', etc), the cyclic structure is necessarily broken. Because the majority of cyclic structures last only briefly ($<1$ ps), the active switching of \suldiox~bonding coordinations appears to break the cyclic structure. Figure \ref{fig:bonding-coordinations} shows that the sum of coordinations that necessarily break cyclic structures account for approximately 39\% and 24\% of the bonding coordinations in the cold and hot system, respectively. Consequently, the distribution of Figure \ref{fig:cyclic-breakdown} also indicates that there are more cycles formed in the hot system (approximately 16\% above the cold). Of the three most encountered bonding coordinations, the hot system shifts population from the ``S'' (a cycle-breaking coordination) to ``SO'' and ``SOO'' (coordinations that allow for bonding cycle formation through the \suldiox). This is relative to the cold system coordination distribution that has equal ``S'' and ``SOO'' populations. The increased temperature appears to cause the \suldiox~to bond in a way that is more conducive to the formation of cyclic structures. This may account for the slightly increased populations of the longer cycle lifespans in the hot system (greater than 4 ps in some cases) compared to the cold system in the cycle lifespan distribution of Figure \ref{fig:cycle-lifespans}.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{cyclic-lifespans-inset-small.png}
		\caption{The lifetime of a cyclic hydrate structure after it forms, and before it distorts in shape due to the movement of the liquid surface molecules, is most often short ($<$ 1 ps). Shown here is a plot of cycle hydrate structure lifetimes after having applied the debouncing procedure (described in the text). Both cold (blue) and hot (red) simulations are graphed. The inset expands the view of the region above 1 ps.}
		\label{fig:cycle-lifespans}
	\end{center}
\end{figure}

>>>>

Quantum MD techniques are the logical follow-up as they accurately reproduce the hydration geometry around the bound \suldiox~molecules, and allow us to examine in detail the specific bonding interactions that occur within the surface hydrates, and in the extended bonding further into the water.\cite{Baer2010} Two parallel studies are performed in this work; one at 300K, and one at the more atmospherically relevant cold 210K. This set of temperatures complements our most recent experimental studies that showed the binding of gaseous \suldiox~to a water surface is greatly enhanced at cold temperatures.\cite{Ota2011} Other experiments by our group developed the picture of \suldiox~adsorption, and showed that \suldiox~surface hydrate complexes form when a water surface is exposed to \suldiox~gas.\cite{Tarbuck2005,Tarbuck2006} Although conclusions regarding the specific nature of those complexes could only be inferred from the experiments, our current computational studies now provide us with insights about the specific microscopic geometries and behaviors of the hydrating complexes.

@@

As noted above, the optimized geometry of the gas phase \suldiox~hydrates suggests cyclic bonding structures. Geometry optimization shows the formation of these cyclic hydrate structures with two or three waters. A different story entirely has the potential to emerge when \suldiox~is placed in a dynamic environment such as in the course of MD simulations of an aqueous surface. Do the cyclic structures also form in the course of a dynamic bonding process on a simulated water surface, where extended hydrating structures influence \suldiox~and water behavior? To study the formation and behavior of cyclic hydrate structures we employ graph theoretical techniques on MD trajectory data. Previous use of graphs in molecular computations were applied to finding stable arrangements of water clusters, ice, hydrogen bonding, extracting topological molecular properties, enumerating complex reaction pathways, and cyclic structure studies.\cite{Anick2002, Huber2007, Radhakrishnan1991, Shi2005, Garcia2004, McDonald1998, Sinanogly1975}

@@

On-the-fly ab initio molecular dynamics simulations were performed with the QUICKSTEP package, which is an implementation of the Gaussian plane wave method using the Kohn-Sham formulation of density functional theory (DFT).\cite{VandeVondele2005} The Kohn-Sham orbitals are expanded using a linear combination of atom-centered Gaussian-type orbital functions. The electronic charge density was described using an auxiliary basis set of plane waves. Energies and forces from on-the-fly simulation sampling of the Born-Oppenheimer surface were calculated for each MD step using the Gaussian TZV2P basis set with a cutoff of 280 Ry, the exchange-correlation functional of Becke, Lee, Yang, and Parr (B3LYP),\cite{LEE1988} and the atomic pseudo-potentials of the Goedecker, Teter, and Hutter type.\cite{Goedecker1996} A simulation timestep of 0.5 fs was used, with a Nos\'{e}-Hoover thermostat set at 210K and 300K for the ``cold'' and ``hot'' simulations, respectively. %These computational parameters were verified to yield a reasonable description of bulk room temperature water when simulating a neat-water system.

Initially, 20 equilibrated boxes of side-lengths 8.545\angs, with 24 randomly packed water molecules were used. Ten of the boxes were used for each of the cold and hot simulations. A sulfur dioxide molecule was randomly placed onto the surface within 2.5\angs~of a water molecule centrally located above the waters in the z-axis. A copy of the initial system cubes were then expanded along one axis (z-axis) to 30\angs. The system energy was minimized through a geometry optimization. Subsequently, the system was equilibrated for 1 ps in canonical ensemble (NVT) conditions. Periodic boundaries were set on the two short axes to form an infinite slab. The equilibrated systems were then simulated for a further 20 ps in the microcanonical ensemble (NVE), with trajectory snapshots recorded every 0.5 fs. The initial 1 ps equilibration trajectory was not included in the final analysis. This simulation process resulted in 40,000 time steps of system trajectory for analysis in each of the hot and cold replicas of the system, for a total of 400,000 timesteps at each temperature.

@@

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{coordination-distributions.png}
		\caption{The \suldiox~bonding coordinations occur with different frequencies at the surface under hot and cold temperature conditions. Shown above is the distribution of bonding coordinations of the cold (blue) and hot (red) \suldiox~over the entire set of trajectories. The values are the percentage of MD timesteps spent in the given bonding coordination.}
		\label{fig:bonding-coordinations}
	\end{center}
\end{figure}

@@

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{cycle-distributions.png}
		\caption{Cyclic hydrate structures form throughout the MD simulations involving the \suldiox~and one or more water molecules. Shown here are the different cycles by number of water (up to 4), and their occurrence rate as a percentage of the total simulation time. The far right part of the plot shows the contributions of the two types of triple-water cycles (see Figure \ref{fig:type-3-varieties}). The cycles had to meet the ``minimally-SO'' bonding criteria (at least one bond to the sulfur and one oxygen bond on the \suldiox) to be used in the analysis.}
		\label{fig:cyclic-breakdown}
	\end{center}
\end{figure}

@@

We know that the \suldiox~bound to a water surface is most likely in the ``SO'' bonding coordination, and is also often taking part in some type of cyclic structure. With this in mind, it becomes interesting to ask: how long does a cyclic hydrate form before breaking to an acyclic hydrate structure? To answer the question of cyclic lifespan, a method was devised to define the lifetime of a cycle. For each MD trajectory, coordinate data was analyzed to determine if a cyclic hydrate structure was formed as described earlier in this manuscript. %A timeline was then produced where each timestep was given a value of 1 or 0 depending on whether a \suldiox~hydrate bonding cycle was present or not, respectively. This resulted in a time-function, $C(t)$, similar in nature to a time-varying digital signal.

%In an electro-mechanical system, a mechanical switch often outputs a noisy signal, full of transients or ``signal bounce'' before settling to a final value. This will appear as a rapid on-off cycling of the signal, and the problem is one of great concern in signal processing. Many mechanical, electrical, and software solutions have been devised to suppress the transients, or ``debounce'' the signal. Analogously, the hydrate bonding cycles formed in the simulated system often undergo a period of time during formation, or before dissolution, where the $C(t)$ function bounces before settling into a final value. The bounce in the function manifests itself in our statistics as a series of rapid switches between cycle presence and cycle absence. Physically, the bond lengths within the cycle (or the cycle-to-be) are fluctuating back and forth across our bond-length criteria. It is thus an artifact of our algorithmic determination of the presence or absence of a bond. Removing this artifact will allow us to examine the longer-time bonding behavior and this is accomplished as follows.

%We wish to find a function based on $C(t)$ that eliminates the brief oscillations during transitions. The resulting function, $f(t)$, will only contain information of whether a hydrate bonding cycle is present or absent, and none of the noisy oscillations of the transitions between the two states. $f(t)$ may then be used to calculate statistics about the cycle lifespans. A representative portion of a cycle time-function, $C(t)$, is plotted for one of the simulated trajectories shown as the dashed black line in Figure \ref{fig:debouncing}. At the far left of the plot the function is in the ``no cycle'' state indicating that a bonding cycle has not been detected, and then switches ``on'' as a cycle is found later in time. The cyclic structure is very dynamic, constantly moving and distorting, so any of the bonds forming the cyclic bonding structure are liable to break and reform quickly. This bouncing between states is manifested in Figure \ref{fig:debouncing} as a series of sharp spikes in the $C(t)$ function lasting less than 10 fs each. $C(t)$ was smoothed using a moving gaussian window function with a 10 fs width, having the effect of disregarding cycle breaks or formations of less than 20 fs duration. The resulting smoothed function, $C_s(t)$, was then cutoff with the following criteria:

%\[
%  f(t) = \left\{ 
%  \begin{array}{l l}
%    0 & \quad C_s(t) < 0.2\\
%    1 & \quad C_s(t) \geq 0.2\\
%  \end{array} \right.
%\]

%where $f(t)$ is the debounced time-function that represents the lifespans of cycles. Figure \ref{fig:debouncing} shows the original time-function of cycle formation and breaking, $C(t)$ (dashed black), the smoothed function, $C_s(t)$ (red), and the final debounced function, $f(t)$ (green).

%\begin{figure}[h!]
%	\begin{center}
%		\includegraphics[scale=1.0]{cycle-debouncing-small.png}
%		\caption{The dynamic environment of the water surface causes rapid breaking and formation of bonds to the \suldiox, and to waters involved in the cyclic hydrating structures. A debouncing procedure was developed (described in the text) to account for the very brief ($<$ 20 fs) breaks to the cyclic structure. A time function, $C(t)$ (dashed black), shows if the bonding distance criteria are met for all members of the cyclic hydrate structure. A gaussian smoothing window was then used to produce $C_s(t)$ (red), which was then used with a cutoff criteria to give the debounced cycle formation time function, $f(t)$ (green).}
%		\label{fig:debouncing}
%	\end{center}
%\end{figure}

The distribution of cycle lifetimes (i.e. contiguous spans of time spent with a fully formed cyclic hydrate) was determined. The distribution of cyclic lifespans is shown in Figure \ref{fig:cycle-lifespans} for the cold (blue) and hot (red) simulations. The most frequent lifespan for both temperature data sets is less than 50 fs. This accounts for the vast majority of cyclic hydrates found (approximately 80\%) indicating that these structures are very transient, and are continuously forming and breaking for very short periods of time. The nature of the water surface, and the very dynamic extended hydrogen bonding network, keeps many of the structures from lasting much longer than a few fs. The difference between hot and cold systems in the $<50$ fs population is less than 3\%, with this trend extending to the longer lifespans as well. The inset of figure \ref{fig:cycle-lifespans} shows an expanded view of the region above 0.2 ps. All of the distribution shows only a $<5$\% difference between cold and hot cyclic lifetimes. Overall, the distribution shows that when cycles form, at both temperatures, they last a similar amount of time. The cold system has more cycles forming than the hot system at the shortest cycle lifetimes ($<10$ fs). Up to 100 fs, the hot temperature cycles show a very slight population increase above the cold temperature cycles.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{cycle-lifespans.png}
		\caption{The lifetime of a cyclic hydrate structure after it forms, and before it distorts in shape due to the movement of the liquid surface molecules, is most often short ($<$ 50 fs). Shown here is a plot of cycle hydrate structure lifetimes. Both cold (blue) and hot (red) simulations are graphed. The inset expands the view of the region above 200 fs.}
		\label{fig:cycle-lifespans}
	\end{center}
\end{figure}

We know from the transition frequency plots of Figure \ref{fig:coordination-transitions} that the bonding coordinations are switching frequently. The \suldiox~is likely forming and breaking bonds with waters external to the cyclic hydrate structures (i.e. not directly involved in the bonds of the cycle). For the longer-lived cycles, the external bonding to the \suldiox~may have little effect on the cyclic hydrate waters. However, any time the \suldiox~switches into an unbound, sulfur-only, or oxygen-only coordination (i.e. ``S'', ``SS'', ``O'', etc), the cyclic structure is necessarily broken. Because the majority of cyclic structures last only briefly ($<1$ ps), the active switching of \suldiox~bonding coordinations appears to break the cyclic structure. Figure \ref{fig:bonding-coordinations} shows that the sum of coordinations that necessarily break cyclic structures account for approximately 54\% and 70\% of the bonding coordinations in the cold and hot system, respectively. Consequently, the distribution of Figure \ref{fig:cyclic-breakdown} also indicates that there are more cycles formed in the hot system (approximately 2.5\% above the cold). Of the most encountered bonding coordinations the hot system has a much higher proportion of unbound coordinations, and a relative decrease of the ``SOO'' coordination. The increased temperature appears to cause the \suldiox~to bond in a way that is more conducive to the formation of cyclic structures. This may account for the slightly increased populations of the longer cycle lifespans in the hot system (greater than 4 ps in some cases) compared to the cold system in the cycle lifespan distribution of Figure \ref{fig:cycle-lifespans}.
